Naki_Main

### 2024-12-06 OldFiles/Naki_EDA_Exploration.qmd
new files

|------------------|--------------------|------------------|------------------|

@@

| Variable | Chi-Square Statistic | Degrees of Freedom (df) | P-Value | Test Type |
|---------------|---------------|---------------|---------------|---------------|

@@

|  |  |  |  |  |
|---------------|---------------|---------------|---------------|---------------|
| Q629R1 | 114.83 | 1 | \< 2.2e-16 | Pearson's Chi-squared test with Yates' correction |

|  |  |  |  |  |
|---------------|---------------|---------------|---------------|---------------|
| Q629R2 | 72.719 | 1 | \< 2.2e-16 | Pearson's Chi-squared test with Yates' correction |

|        |         |     |        |                                                   |
|---------------|---------------|---------------|---------------|---------------|
| Q629R3 | 0.41214 | 1   | 0.5209 | Pearson's Chi-squared test with Yates' correction |

|        |        |     |        |                                                   |
|---------------|---------------|---------------|---------------|---------------|
| Q629R4 | 1.3169 | 1   | 0.2511 | Pearson's Chi-squared test with Yates' correction |

|        |       |     |           |                                                   |
|---------------|---------------|---------------|---------------|---------------|
| Q629R5 | 11.48 | 1   | 0.0007036 | Pearson's Chi-squared test with Yates' correction |

|  |  |  |  |  |
|---------------|---------------|---------------|---------------|---------------|
| Q629R6 | 20.113 | 1 | 7.301e-06 | Pearson's Chi-squared test with Yates' correction |

|        |        |     |        |                                                   |
|---------------|---------------|---------------|---------------|---------------|
| Q629R7 | 2.6287 | 1   | 0.1049 | Pearson's Chi-squared test with Yates' correction |

>>>>

|----|----|----|----|

@@

| Variable | Chi-Square Statistic | Degrees of Freedom (df) | P-Value | Test Type |
|----------|----------------------|-------------------------|---------|-----------|

@@

|  |  |  |  |  |
|----|----|----|----|----|
| Q629R1 | 114.83 | 1 | \< 2.2e-16 | Pearson's Chi-squared test with Yates' correction |

|  |  |  |  |  |
|----|----|----|----|----|
| Q629R2 | 72.719 | 1 | \< 2.2e-16 | Pearson's Chi-squared test with Yates' correction |

|        |         |     |        |                                                   |
|--------|---------|-----|--------|---------------------------------------------------|
| Q629R3 | 0.41214 | 1   | 0.5209 | Pearson's Chi-squared test with Yates' correction |

|        |        |     |        |                                                   |
|--------|--------|-----|--------|---------------------------------------------------|
| Q629R4 | 1.3169 | 1   | 0.2511 | Pearson's Chi-squared test with Yates' correction |

|        |       |     |           |                                                   |
|--------|-------|-----|-----------|---------------------------------------------------|
| Q629R5 | 11.48 | 1   | 0.0007036 | Pearson's Chi-squared test with Yates' correction |

|  |  |  |  |  |
|----|----|----|----|----|
| Q629R6 | 20.113 | 1 | 7.301e-06 | Pearson's Chi-squared test with Yates' correction |

|        |        |     |        |                                                   |
|--------|--------|-----|--------|---------------------------------------------------|
| Q629R7 | 2.6287 | 1   | 0.1049 | Pearson's Chi-squared test with Yates' correction |

### 2024-12-06 OldFiles/Naki_Main1.qmd
new files

@@

## Extracting relevant columns
```{r}
# Define the required columns
required_cols <- c(
  # Outcome Variable: Stopping discrimination against racial/ethnic minorities
  "Q1R5",

@@

  # Perceptions of discrimination
  "Q619_Q626R1",     # Perceptions of discrimination against Whites
  "Q619_Q626R2",     # Perceptions of discrimination against Blacks
  "Q619_Q626R3",     # Perceptions of discrimination against Asian Americans
  "Q619_Q626R4",     # Perceptions of discrimination against Native Americans
  "Q619_Q626R5",     # Perceptions of discrimination against Immigrants
  "Q619_Q626R6",     # Perceptions of discrimination against Latinos
  "Q619_Q626R7",     # Perceptions of discrimination against GaysAndLesbians
  "Q619_Q626R8"      # Perceptions of discrimination against Muslims
)
```

## Cleaning and factoring data for all columns
```{r}
# Create a new dataframe with selected variables
dataset_new <- dataset[, required_cols]

# Recode the outcome variable 'Q1R5' for readability
dataset_new$Q1R5 <- factor(
  ifelse(
    dataset_new$Q1R5 == "(0) NO TO: Stopping discrimination against racial/ethnic minorities", 
    "Not important",
    ifelse(dataset_new$Q1R5 == "(1) Stopping discrimination against racial/ethnic minorities", 
           "Important", 
           "Not specified")
  ),
  levels = c("Not important", "Important", "Not specified") # Define factor levels
)

# Define the columns for perceptions of discrimination
perception_cols <- c(
  "Q619_Q626R1",  # Perceptions of discrimination against Whites
  "Q619_Q626R2",  # Perceptions of discrimination against Blacks
  "Q619_Q626R3",  # Perceptions of discrimination against Asian Americans
  "Q619_Q626R4",  # Perceptions of discrimination against Native Americans
  "Q619_Q626R5",  # Perceptions of discrimination against Immigrants
  "Q619_Q626R6",  # Perceptions of discrimination against Latinos
  "Q619_Q626R7",  # Perceptions of discrimination against GaysAndLesbians
  "Q619_Q626R8",  # Perceptions of discrimination against Muslims
  "Q633"          # Discrimination impact on life satisfaction
)

# Recode 'Don' as 'Don't know' and ensure columns remain factors
dataset_new[perception_cols] <- lapply(dataset_new[perception_cols], function(x) {
  x <- as.character(x)  # Temporarily convert to character for recoding
  x <- ifelse(x == "(5) Don", "(5) Don't know", x)  # Recode 'Don' to 'Don't know'
  factor(x)  # Convert back to factor
})

# Verify the changes in perception columns
head(dataset_new[perception_cols])

# Define required columns to convert to factors (if not already defined)
# Assuming `required_cols` is previously defined; otherwise, specify explicitly
# e.g., required_cols <- c("Q1R5", "Q619_Q626R1", ...)
dataset_new[required_cols] <- lapply(dataset_new[required_cols], as.factor)

@@

## Consolidating Religion to one column
```{r}
# Convert individual religion columns to numeric
for (i in 1:15) {
  col_name <- paste0("Q58R", i, "_numeric")
  dataset_new[[col_name]] <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", dataset_new[[paste0("Q58R", i)]]))
}

@@

# Take a look at the updated dataset
head(dataset_clean)
```

@@

```{r}
library(vcd)
library(tidyverse)
library(corrplot)
```

@@

        data = dataset_clean, family = "binomial")
```

### Check VIF values
```{r}
# Check multicollinearity using VIF
vif(model)
```

```{r}
## # Extract odds ratios
## odds_ratios <- exp(coef(model))
## 
## # Get confidence intervals for coefficients
## conf_intervals <- exp(confint(model))
## 
## # Extract p-values
## p_values <- summary(model)$coefficients[, 4]
## 
## # Combine everything into a single data frame
## results_model <- data.frame(
##   OddsRatio = odds_ratios,
##   LowerCI = conf_intervals[, 1],
##   UpperCI = conf_intervals[, 2],
##   PValue = p_values
## )
## 
## # Display results
## results_model
```

## Perform stepwise selection to select best parameters
```{r}
# Perform stepwise selection
model_stepwise <- step(model, 
                       direction = "both",
                       trace = 0)
```

### Check VIF Values
```{r}
# Check multicollinearity using VIF
vif(model_stepwise)
```

## Extract Odds for each variable in the stepwise model
```{r}
# Extract odds ratios
odds_ratios <- exp(coef(model_stepwise))

@@

# Display results
results_model_stepwise
```

## Plot showing odds
```{r}
# Filter out only significant variables (p-value < 0.05)
significant_data <- results_model_stepwise[results_model_stepwise$PValue < 0.051, ]

# Create the plot with inverted axes and CI shown as error bars
ggplot(significant_data, aes(y = rownames(significant_data), x = OddsRatio)) +
  geom_point(size = 4, color = "blue") +  # Plot the odds ratio as points
  geom_errorbarh(aes(xmin = LowerCI, xmax = UpperCI), height = 0.2, color = "black") +  # Add horizontal error bars
  labs(title = "Significant Variables with Confidence Intervals (Odds Ratio)",
       x = "Odds Ratio",
       y = "Variables") +
  theme(axis.text.y = element_text(angle = 0, hjust = 1))  # Keep y-axis labels horizontal
  theme(plot.margin = margin(1, 2, 1, 1, "cm"))  # Adjust margins for larger size

# Save the plot with larger dimensions
ggsave("significant_variables_plot.png", width = 12, height = 8)
```

```{r}
# Filter for significant variables (p-value < 0.05)
significant_data <- model_stepwise |> 
  filter(PValue < 0.05)

# Prepare data for plotting
# You might need to adjust this section to ensure the data is in long format with appropriate variable names
plot_data <- significant_data |>
  select(Variable, OddsRatio) |>
  mutate(Variable = factor(Variable, levels = unique(Variable)))

# Create the boxplot
ggplot(plot_data, aes(x = Variable, y = OddsRatio)) +
  geom_boxplot() +
  labs(title = "Boxplot of Significant Variables (Odds Ratio)",
       x = "Variables",
       y = "Odds Ratio") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

>>>>

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

The `echo: false` option disables the printing of code (only output is displayed).

@@

## Cleaning and factoring all columns in the df

```{r}
# Define the required columns
required_cols <- c(
  # Outcome Variable: Stopping discrimination against racial/ethnic minorities
  "Q1R5",

@@

  # Perceptions of discrimination
  "Q619_Q626R1",     # Perceptions of discrimination against Whites
  "Q619_Q626R2",     # Perceptions of discrimination against Blacks
  "Q619_Q626R3",     # Perceptions of discrimination against Asian Americans
  "Q619_Q626R4",     # Perceptions of discrimination against Native Americans
  "Q619_Q626R5",     # Perceptions of discrimination against Immigrants
  "Q619_Q626R6",     # Perceptions of discrimination against Latinos
  "Q619_Q626R7",     # Perceptions of discrimination against GaysAndLesbians
  "Q619_Q626R8"      # Perceptions of discrimination against Muslims
)

# Create a new dataframe with selected variables
dataset_new <- dataset[, required_cols]

# Recode the outcome variable 'Q1R5' for readability
dataset_new$Q1R5 <- ifelse(
  dataset_new$Q1R5 == "(0) NO TO: Stopping discrimination against racial/ethnic minorities", 
  "Not important",
  ifelse(dataset_new$Q1R5 == "(1) Stopping discrimination against racial/ethnic minorities", 
         "Important", 
         "Not specified")
)

# Define the columns for perceptions of discrimination
perception_cols <- c(
  "Q619_Q626R1",  # Perceptions of discrimination against Whites
  "Q619_Q626R2",  # Perceptions of discrimination against Blacks
  "Q619_Q626R3",  # Perceptions of discrimination against Asian Americans
  "Q619_Q626R4",  # Perceptions of discrimination against Native Americans
  "Q619_Q626R5",  # Perceptions of discrimination against Immigrants
  "Q619_Q626R6",  # Perceptions of discrimination against Latinos
  "Q619_Q626R7",  # Perceptions of discrimination against GaysAndLesbians
  "Q619_Q626R8",  # Perceptions of discrimination against Muslims

  "Q633"          # Discrimination impact on life satisfaction
)

# Recode 'Don' as 'Don't know'
dataset_new[perception_cols] <- lapply(dataset_new[perception_cols], function(x) {
  x <- ifelse(x == "Don", "Don't know", x)  # Recode 'Don' to 'Don't know'
  return(x)
})

# Verify the changes
head(dataset_new[perception_cols])

# Convert all required columns to factors
dataset_new[required_cols] <- lapply(dataset_new[required_cols], as.factor)

@@

This code prepares the dataset for further analysis by selecting and cleaning up the relevant variables, ensuring they are in a suitable format (i.e., factors for categorical data).

**Q633 & Q619_Q626R1 - Q619_Q626R8 Category Key**

-   1 - A lot

-   2 - Some

-   3 - A little

-   4 - None at all

-   5 - Don't Know

    ## Changing factor key

    ```{r}
    # Define the mapping from numeric values to the desired labels
    perception_labels <- c("A lot", "Some", "A little", "None at all", "Don't know")

    # Convert the factored perception columns to the correct labels
    dataset_new[perception_cols] <- lapply(dataset_new[perception_cols], function(x) {
      # Convert the numeric factors (1:5) to the corresponding perception labels
      x <- factor(x, levels = 1:5, labels = perception_labels)
      return(x)
    })

    # Verify the changes
    head(dataset_new[perception_cols])

    # Check the structure
    str(dataset_new[perception_cols])

    ## Consolidating Religion to one column

```{r}
# Convert individual religion columns to numeric
for (i in 1:15) {
  col_name <- paste0("Q58R", i, "_numeric")
  dataset_new[[col_name]] <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", dataset_new[[paste0("Q58R", i)]]))
}

@@

# Take a look at the updated dataset
head(dataset_clean)

@@

@@

        data = dataset_clean, family = "binomial")

# summary(model)
```

```{r}
# Extract odds ratios
#odds_ratios <- exp(coef(model))

# Get confidence intervals for coefficients
#conf_intervals <- exp(confint(model))

# Extract p-values
#p_values <- summary(model)$coefficients[, 4]

# Combine everything into a single data frame
#results_model <- data.frame(
#  OddsRatio = odds_ratios,
#  LowerCI = conf_intervals[, 1],
#  UpperCI = conf_intervals[, 2],
#  PValue = p_values
#)

# Display results
# results_model

# Check multicollinearity using VIF
# vif(model)
```

## Model Stepwise

```{r}
# Perform stepwise selection
model_stepwise <- step(model, 
                       direction = "both",
                       trace = 0)
```

```{r}
# Extract odds ratios
odds_ratios <- exp(coef(model_stepwise))

@@

# Display results
results_model_stepwise

# Check multicollinearity using VIF
# vif(model_stepwise)
```

```{r}
predicted_classes <- ifelse(predicted_probabilities > 0.7, "Not important", "Important")
actual_classes <- dataset_clean$ImportanceOfStoppingDiscriminationAgainstMinorities
```

```{r}
#predicted_classes <- ifelse(predicted_probabilities > 0.5, 1, 0)
confusion_matrix <- table(Predicted = predicted_classes, Actual = actual_classes)
print(confusion_matrix)

@@

```{r}
library(caret)
confusion <- confusionMatrix(factor(predicted_classes), factor(actual_classes))
print(confusion)
```

### 2024-12-06 OldFiles/Naki_Main_Old.qmd
new files

---
title: "Naki_Main"
format: html
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

The `echo: false` option disables the printing of code (only output is displayed).

```{r}

# Load the packages
library(ggplot2)
library(broom)
library(dplyr)
library(car)
library(caret)
```

# Introduction

@@

# Preliminary Cleaning

## Loading the Dataset

```{r}
file_path <- "/Users/nakiyahdhariwala/StatsProject/39096-0001-Data.rda"

load(file_path)          # Load the RDA file
ls()                     # List objects loaded into the environment
dataset <- da39096.0001  # Access the dataset

head(dataset)
ncol(dataset)
nrow(dataset)
```

@@

# Plot the pie chart with custom labels
pie(proportions,
    labels = pie_labels,
    main = "Q1R5: Stopping discrimination against racial/ethnic minorities",
    col = c("lightblue", "salmon"))
```

The outcome variable Q1R5 (Stopping discrimination against racial/ethnic minorities) for this Research Question is a binary variable which captures whether respondents view stopping racial and ethnic discrimination as an important issue (i.e., Important OR Not Important). Therefore, we will make use Logistic regression as it is suitable for modeling binary outcome variables and allows for the inclusion of both predictor and control variables. While the distribution is skewed (74.25% say "No"), logistic regression is robust to imbalance in the outcome variable. However, extreme imbalance (e.g., 95%-5%) might require additional techniques like resampling or penalized logistic regression.

@@

## Cleaning and factoring all columns in the df

```{r}
# Define the required columns
required_cols <- c(
  # Outcome Variable: Stopping discrimination against racial/ethnic minorities
  "Q1R5",

  # Demographic Variables
  "S2_RACE_PRIME",   # Primary Race
  "S3",              # SexualOrientation
  "S3B",             # Gender
  "S5_AGE",          # Age categories
  "S13",             # Highest level of education
  "S14",             # Community lived in

@@

  # Perceptions of discrimination
  "Q619_Q626R1",     # Perceptions of discrimination against Whites
  "Q619_Q626R2",     # Perceptions of discrimination against Blacks
  "Q619_Q626R3",     # Perceptions of discrimination against Asian Americans
  "Q619_Q626R4",     # Perceptions of discrimination against Native Americans
  "Q619_Q626R5",     # Perceptions of discrimination against Immigrants
  "Q619_Q626R6",     # Perceptions of discrimination against Latinos
  "Q619_Q626R7",     # Perceptions of discrimination against GaysAndLesbians
  "Q619_Q626R8"      # Perceptions of discrimination against Muslims
)

@@

# Define the columns for perceptions of discrimination
perception_cols <- c(
  "Q619_Q626R1",  # Perceptions of discrimination against Whites
  "Q619_Q626R2",  # Perceptions of discrimination against Blacks
  "Q619_Q626R3",  # Perceptions of discrimination against Asian Americans
  "Q619_Q626R4",  # Perceptions of discrimination against Native Americans
  "Q619_Q626R5",  # Perceptions of discrimination against Immigrants
  "Q619_Q626R6",  # Perceptions of discrimination against Latinos
  "Q619_Q626R7",  # Perceptions of discrimination against GaysAndLesbians
  "Q619_Q626R8",  # Perceptions of discrimination against Muslims

  "Q633"          # Discrimination impact on life satisfaction
)

# Recode 'Don' as 'Don't know'
dataset_new[perception_cols] <- lapply(dataset_new[perception_cols], function(x) {
  x <- ifelse(x == "Don", "Don't know", x)  # Recode 'Don' to 'Don't know'
  return(x)
})

# Verify the changes
head(dataset_new[perception_cols])

@@

# Display the first few rows to check the changes
head(dataset_new)
```

This code prepares the dataset for further analysis by selecting and cleaning up the relevant variables, ensuring they are in a suitable format (i.e., factors for categorical data).

**Q633 & Q619_Q626R1 - Q619_Q626R8 Category Key**

-   1 - A lot

-   2 - Some

-   3 - A little

-   4 - None at all

-   5 - Don't Know

    ## Changing factor key

    ```{r}
    # Define the mapping from numeric values to the desired labels
    perception_labels <- c("A lot", "Some", "A little", "None at all", "Don't know")

    # Convert the factored perception columns to the correct labels
    dataset_new[perception_cols] <- lapply(dataset_new[perception_cols], function(x) {
      # Convert the numeric factors (1:5) to the corresponding perception labels
      x <- factor(x, levels = 1:5, labels = perception_labels)
      return(x)
    })

    # Verify the changes
    head(dataset_new[perception_cols])

    # Check the structure
    str(dataset_new[perception_cols])

    ## Consolidating Religion to one column

# Create a new Religion variable
dataset_new$Religion <- "None"

# Assign religion based on the numeric values of each Q58R column
religion_mapping <- c(
  "Q58R1_numeric" = "Catholic", 
  "Q58R2_numeric" = "Protestant", 
  "Q58R3_numeric" = "Other Christian",
  "Q58R4_numeric" = "Jewish", 
  "Q58R5_numeric" = "Muslim", 
  "Q58R6_numeric" = "Hindu", 
  "Q58R7_numeric" = "Buddhist", 
  "Q58R8_numeric" = "Mormon/LDS", 
  "Q58R9_numeric" = "Folk Religion", 
  "Q58R10_numeric" = "Daoist", 
  "Q58R11_numeric" = "Shamanism", 
  "Q58R12_numeric" = "Ancestral Religion", 
  "Q58R13_numeric" = "Spiritual but not religious", 
  "Q58R14_numeric" = "Atheist", 
  "Q58R15_numeric" = "Agnostic"
)

# Loop through the mapping to assign values
for (col in names(religion_mapping)) {
  dataset_new$Religion[dataset_new[[col]] == 1] <- religion_mapping[[col]]
}

# Convert Religion to a factor with specific levels
dataset_new$Religion <- factor(dataset_new$Religion,
                               levels = c("Catholic", "Protestant", "Other Christian", "Jewish", "Muslim",
                                          "Hindu", "Buddhist", "Mormon/LDS", "Folk Religion",
                                          "Daoist", "Shamanism", "Ancestral Religion",
                                          "Spiritual but not religious", "Atheist", "Agnostic", 
                                          "Multiple Religions", "None"))

# Remove redundant religion columns (Q58R1 to Q58R15 and their numeric versions)
redundant_cols <- c(
  paste0("Q58R", 1:15), paste0("Q58R", 1:15, "_numeric")
)

# Take a look at the updated dataset
head(dataset_clean)

@@

@@

# Convert to a data frame
missing_summary <- data.frame(Variable = names(missing_proportions),
                              Missing_Proportion = round(missing_proportions, 4))

@@

head(dataset_clean)
```

## Rename the remaining column headers for ease of understanding
```{r}
colnames(dataset_clean)
colnames(dataset_clean) <- c(
  "ImportanceOfStoppingDiscriminationAgainstMinorities",
  "PrimaryRace", 
  "Orientation", 
  "Gender", 
  "Age", 
  "HighestLevelOfEducation", 
  "CommunityLivedIn", 
  "TreatedUnfairlyOrDiscriminatedAgainst", 
  "DiscriminationImpactOnLifeSatisfaction",
  "PerceptionsOfDiscriminationAgainstWhites",
  "PerceptionsOfDiscriminationAgainstBlacks", 
  "PerceptionsOfDiscriminationAgainstAsianAmericans", 
  "PerceptionsOfDiscriminationAgainstNativeAmericans", 
  "PerceptionsOfDiscriminationAgainstImmigrants", 
  "PerceptionsOfDiscriminationAgainstLatinos", 
  "PerceptionsOfDiscriminationAgainstGaysAndLesbians",
  "PerceptionsOfDiscriminationAgainstMuslims",
  "Religion"
)

@@

# Binomial Logistic Regression

## Model

```{r}
model <- glm(ImportanceOfStoppingDiscriminationAgainstMinorities ~ PrimaryRace +
               Orientation + Gender + Age + 
               HighestLevelOfEducation + CommunityLivedIn +
               TreatedUnfairlyOrDiscriminatedAgainst + 
               DiscriminationImpactOnLifeSatisfaction + 
               PerceptionsOfDiscriminationAgainstWhites + 
               PerceptionsOfDiscriminationAgainstBlacks +
               PerceptionsOfDiscriminationAgainstAsianAmericans + 
               PerceptionsOfDiscriminationAgainstNativeAmericans + 
               PerceptionsOfDiscriminationAgainstImmigrants + 
               PerceptionsOfDiscriminationAgainstLatinos + 
               PerceptionsOfDiscriminationAgainstGaysAndLesbians + 
               PerceptionsOfDiscriminationAgainstMuslims + 
               Religion,

        data = dataset_clean, family = "binomial")

# summary(model)
```

```{r}
# Extract odds ratios
#odds_ratios <- exp(coef(model))

# Get confidence intervals for coefficients
#conf_intervals <- exp(confint(model))

# Extract p-values
#p_values <- summary(model)$coefficients[, 4]

# Combine everything into a single data frame
#results_model <- data.frame(
#  OddsRatio = odds_ratios,
#  LowerCI = conf_intervals[, 1],
#  UpperCI = conf_intervals[, 2],
#  PValue = p_values
#)

# Display results
# results_model

# Check multicollinearity using VIF
# vif(model)
```

## Model Stepwise

```{r}
# Perform stepwise selection
model_stepwise <- step(model, 
                       direction = "both",
                       trace = 0)
```

```{r}
# Extract odds ratios
odds_ratios <- exp(coef(model_stepwise))

# Get confidence intervals for coefficients
conf_intervals <- exp(confint(model_stepwise))

# Extract p-values
p_values <- summary(model_stepwise)$coefficients[, 4]

# Combine everything into a single data frame
results_model_stepwise <- data.frame(
  OddsRatio = odds_ratios,
  LowerCI = conf_intervals[, 1],
  UpperCI = conf_intervals[, 2],
  PValue = p_values
)

# Display results
results_model_stepwise

# Check multicollinearity using VIF
# vif(model_stepwise)
```

```{r}
predicted_classes <- ifelse(predicted_probabilities > 0.7, "Not important", "Important")
actual_classes <- dataset_clean$ImportanceOfStoppingDiscriminationAgainstMinorities
```

```{r}
#predicted_classes <- ifelse(predicted_probabilities > 0.5, 1, 0)
confusion_matrix <- table(Predicted = predicted_classes, Actual = actual_classes)
print(confusion_matrix)

```{r}
library(caret)
confusion <- confusionMatrix(factor(predicted_classes), factor(actual_classes))
print(confusion)
```

```{r}

>>>>

```{r}
# Load the packages
library(ggplot2)
library(broom)
library(dplyr)
library(car)
library(caret)
```

# Loading dataset

```{r}
# Set the file path
file_path <- "/Users/nakiyahdhariwala/StatsProject/39096-0001-Data.rda"

# Load the RDA file
load(file_path)

# List objects loaded into the environment
ls()

# Access the dataset (replace `da39096.0001` with the actual name found in `ls()`)
dataset <- da39096.0001  # Use the correct object name
```

```{r}
head(dataset)
ncol(dataset)
nrow(dataset)
```

@@

@@

# Plot the pie chart with custom labels
pie(proportions,
    labels = pie_labels,
    main = "Q1R5: Stopping discrimination against racial/ethnic minorities",
    col = c("lightblue", "salmon"))

The outcome variable Q1R5 (Stopping discrimination against racial/ethnic minorities) for this Research Question is a binary variable which captures whether respondents view stopping racial and ethnic discrimination as an important issue (i.e., 1 = Important, 0 = Not Important). Therefore, we will make use Logistic regression as it is suitable for modeling binary outcome variables and allows for the inclusion of both predictor and control variables. While the distribution is skewed (74.25% say "No"), logistic regression is robust to imbalance in the outcome variable. However, extreme imbalance (e.g., 95%-5%) might require additional techniques like resampling or penalized logistic regression.

@@

```{r}
# Define the required columns
required_cols <- c(
  # Outcome Variable: Stopping discrimination against racial/ethnic minorities (Binary: 0 = Not important, 1 = Important)
  "Q1R5",

  # Demographic Variables
  "S2_RACE_PRIME",   # Primary Race
  "S3",              # How you think of yourself
  "S3B",             # Gender
  "S5_AGE",          # Age categories
  "S13",             # Highest level of education
  "S14",             # Community lived in

@@

  # Perceptions of discrimination
  "Q619_Q626R2",     # Perceptions of discrimination against Blacks
  "Q619_Q626R3",     # Perceptions of discrimination against Asian Americans
  "Q619_Q626R4",     # Perceptions of discrimination against Native Americans
  "Q619_Q626R5",     # Perceptions of discrimination against Immigrants
  "Q619_Q626R6"      # Perceptions of discrimination against Latinos
)

@@

@@

# Display the first few rows to check the changes
head(dataset_new)

@@

# Consolidating Religion into one column

# Create a new Religion variable
dataset_new$Religion <- "None"
dataset_new$Religion[dataset_new$Q58R1_numeric == 1] <- "Catholic"
dataset_new$Religion[dataset_new$Q58R2_numeric == 1] <- "Protestant"
dataset_new$Religion[dataset_new$Q58R3_numeric == 1] <- "Other Christian"
dataset_new$Religion[dataset_new$Q58R4_numeric == 1] <- "Jewish"
dataset_new$Religion[dataset_new$Q58R5_numeric == 1] <- "Muslim"
dataset_new$Religion[dataset_new$Q58R6_numeric == 1] <- "Hindu"
dataset_new$Religion[dataset_new$Q58R7_numeric == 1] <- "Buddhist"
dataset_new$Religion[dataset_new$Q58R8_numeric == 1] <- "Mormon/LDS"
dataset_new$Religion[dataset_new$Q58R9_numeric == 1] <- "Folk Religion"
dataset_new$Religion[dataset_new$Q58R10_numeric == 1] <- "Daoist"
dataset_new$Religion[dataset_new$Q58R11_numeric == 1] <- "Shamanism"
dataset_new$Religion[dataset_new$Q58R12_numeric == 1] <- "Ancestral Religion"
dataset_new$Religion[dataset_new$Q58R13_numeric == 1] <- "Spiritual but not religious"
dataset_new$Religion[dataset_new$Q58R14_numeric == 1] <- "Atheist"
dataset_new$Religion[dataset_new$Q58R15_numeric == 1] <- "Agnostic"

# Check if an individual reported multiple religious affiliations 
#dataset_new$multiple_religions <- rowSums(dataset_new[, grep("_numeric", names(dataset_new))] == 1)
#check <- table(dataset_new$multiple_religions)

# Convert Religion to a factor
dataset_new$Religion <- factor(dataset_new$Religion,
                               levels = c("Catholic", "Protestant", "Other Christian", "Jewish", "Muslim",
                                          "Hindu", "Buddhist", "Mormon/LDS", "Folk Religion",
                                          "Daoist", "Shamanism", "Ancestral Religion",
                                          "Spiritual but not religious", "Atheist", "Agnostic", 
                                          "Multiple Religions", "None"))

# Remove redundant religion columns (Q58R1 to Q58R15 and their numeric versions)
redundant_cols <- c(
  "Q58R1", "Q58R2", "Q58R3", "Q58R4", "Q58R5", "Q58R6", "Q58R7", "Q58R8", 
  "Q58R9", "Q58R10", "Q58R11", "Q58R12", "Q58R13", "Q58R14", "Q58R15",
  "Q58R1_numeric", "Q58R2_numeric", "Q58R3_numeric", "Q58R4_numeric", 
  "Q58R5_numeric", "Q58R6_numeric", "Q58R7_numeric", "Q58R8_numeric", 
  "Q58R9_numeric", "Q58R10_numeric", "Q58R11_numeric", "Q58R12_numeric", 
  "Q58R13_numeric", "Q58R14_numeric", "Q58R15_numeric"
)

# Take a look at the updated dataset
head(dataset_clean)
```

@@

# Convert to a data frame for better readability
missing_summary <- data.frame(Variable = names(missing_proportions),
                              Missing_Proportion = round(missing_proportions, 4))

@@

head(dataset_clean)
#colnames(dataset_clean)
```

# Rename the remaining column headers for ease of understanding

```{r}
colnames(dataset_clean)
colnames(dataset_clean) <- c(
  "ImportanceOfStoppingDiscriminationAgainstMinorities",
  "PrimaryRace", 
  "Orientation", 
  "Gender", 
  "Age", 
  "HighestLevelOfEducation", 
  "CommunityLivedIn", 
  "TreatedUnfairlyOrDiscriminatedAgainst", 
  "DiscriminationImpactOnLifeSatisfaction", 
  "PerceptionsOfDiscriminationAgainstBlacks", 
  "PerceptionsOfDiscriminationAgainstAsianAmericans", 
  "PerceptionsOfDiscriminationAgainstNativeAmericans", 
  "PerceptionsOfDiscriminationAgainstImmigrants", 
  "PerceptionsOfDiscriminationAgainstLatinos", 
  "Religion"
)
colnames(dataset_clean)
```

```{r}
# Fit a logistic regression model
model <- glm(Q1R5 ~ S2_RACE_PRIME + # Primary Race
               S3 +                 # Orientation
               S3B +                # Gender
               S5_AGE +             # Age
               S13 +                # Highest level of education
               S14 +                # Community lived in
               Q627 +               # Treated unfairly or discriminated against
               Q633 +               # Discrimination impact on life satisfaction
               Q619_Q626R2 +        # Perceptions of discrimination against Blacks
               Q619_Q626R3 +        # Perceptions of discrimination against Asian Americans
               Q619_Q626R4 +        # Perceptions of discrimination against Native Americans
               Q619_Q626R5 +        # Perceptions of discrimination against Immigrants
               Q619_Q626R6 +        # Perceptions of discrimination against Latinos
               Religion +
               S3B * Q627,
             data = dataset_clean, family = binomial)

#summary(model)
```

```{r}
# Perform stepwise selection
stepwise_model <- step(model, direction = "both", trace = 1)
```

```{r}
# New model based on stepwise regression
model_final <- glm(Q1R5 ~ S2_RACE_PRIME + # Primary Race
#REMOVED               S3 +                 # Orientation
               S3B +                # Gender
               S5_AGE +             # Age
               S13 +                # Highest level of education
#REMOVED               S14 +                # Community lived in
               Q627 +               # Treated unfairly or discriminated against
               Q633 +               # Discrimination impact on life satisfaction
               Q619_Q626R2 +        # Perceptions of discrimination against Blacks
               Q619_Q626R3 +        # Perceptions of discrimination against Asian Americans
               Q619_Q626R4 +        # Perceptions of discrimination against Native Americans
               Q619_Q626R5,        # Perceptions of discrimination against Immigrants
#REMOVED               Q619_Q626R6 +        # Perceptions of discrimination against Latinos
#REMOVED               Religion +
#REMOVED              S2_RACE_PRIME * S3B,

            data = dataset_clean, family = binomial)

summary(model_final)
```

# Diagnostic Checks

```{r}
# Calculate VIFs
vif_results <- vif(model)  # Uses type = "terms" automatically if predictors are weighted or non-linear
vif_results_model_final <- vif(model_final)

print(vif_results)
print(vif_results_model_final)
```

```{r}
plot(residuals(model, type = "deviance"))
plot(residuals(model_final, type = "deviance"))

influencePlot(model)
influencePlot(model_final)
```

```{r}
levels(as.factor(predictions))
levels(as.factor(dataset_clean$Q1R5))
```

```{r}
levels(as.factor(testData$Q1R5))
predictions <- predict(model_final, testData)

set.seed(123)  # For reproducibility
trainIndex <- createDataPartition(dataset_clean$Q1R5, p = 0.8, list = FALSE)
trainData <- dataset_clean[trainIndex, ]
testData <- dataset_clean[-trainIndex, ]

levels(as.factor(predictions))
levels(as.factor(testData$Q1R5))

confusionMatrix(predictions, testData$Q1R5)

@@
